---
title: "R Notebook"
output: html_notebook
---

Exploring library wifi data.

Read it in. This data not to be store on GitHub or otherwise made publically available.

```{r message=FALSE}
library(tidyverse)
wifi <- readRDS("/Users/jcf2d/Box Sync/WiFi_data/wifi.rds")
wifi$location <- factor(wifi$location)
```

Investigating records that have all `Department`, `Display_Department`, `MBU`, `Registrar_School`, and `uvaPersonIAMAffiliation` all missing. Why is that? Should we drop them? Are they occurring at certain locations?


Add indicator that identifies records with all vars of interest missing
```{r}
vars <- names(wifi)[-(1:7)]
wifi <- mutate(wifi, allNA = apply(wifi[vars], 1, function(x)all(is.na(x))))
```

Quick count

```{r}
scales::percent(mean(wifi$allNA))
scales::comma(sum(wifi$allNA))
```

Now by location

```{r}
wifi %>% 
  group_by(location) %>% 
  summarise(sum_na = sum(allNA),
            mean_na = mean(allNA))
```

Create two data frames: one with records where all vars of interest missing, and one with no such records

```{r}
wifi_allNA <- wifi %>% filter(allNA)
wifi_noNA <- wifi %>% filter(!allNA)
# confirm they add up
nrow(wifi_noNA) + nrow(wifi_allNA) == nrow(wifi)
```

Are there mac addresses in both data frames?

```{r}
length(intersect(wifi_noNA$mac, wifi_allNA$mac))
which_mac <- intersect(wifi_noNA$mac, wifi_allNA$mac)
```

Yes there is. Let's explore. Is there some pattern to them?

```{r}
wifi2 <- filter(wifi, mac %in% which_mac)
map(wifi2[vars], ~ table(.x))
```

Apparently not? Maybe some pattern by location?

```{r}
wifi2 %>% 
  group_by(location) %>% 
  summarise(sum_na = sum(allNA),
            mean_na = mean(allNA))
```

Seems to happen mostly in Harrison and Clemons. 

It appears that records having `Department`, `Display_Department`, `MBU`, `Registrar_School` and `uvaPersonIAMAffiliation` all missing are not always that way. For example, see mac 3673 (user 12800). Has only one instance of allNA at Clemons, but has all NA at Alderman.

```{r}
wifi2 %>% 
  filter(mac == 3673) %>% 
  group_by(location) %>% 
  summarise(sum_na = sum(allNA),
            mean_na = mean(allNA))
```

Does this pattern happen all the time?

```{r}
wifi2 %>% 
  group_by(mac, location) %>% 
  summarise(mean_na = round(mean(allNA), 2)) %>% 
  as.data.frame()
```

Apparently not. See mac 19698. Has all missing at Clemons and none at Alderman.

```{r}
# mac in wifi_noNA not in wifi_allNA
length(setdiff(wifi_noNA$mac, wifi_allNA$mac))

# mac in wifi_allNA not in wifi
length(setdiff(wifi_allNA$mac, wifi_noNA$mac))
```

Appears to be over 16,500 unique macs that always have all `vars` missing

```{r}
n_distinct(wifi$mac)
length(setdiff(wifi_allNA$mac, wifi_noNA$mac))/ n_distinct(wifi$mac)
```

That's about 26% of the unique macs.